---
title: "dataQAQC"
author: "Alene Onion"
date: "December 29, 2018"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#file.copy("sections/images", ".", recursive = TRUE)
```

##QAQC overview

The components of our quality review are:

* Laboratory QAQC
    + Checking that laboratory analyses met all internal QAQC checks  
* Accuracy
    + Assessing the bias of a method in a given sample matrix Using a sample matrix spike   
* Precision   
    + Assessing method precision using duplicate samples analyzed from the same date/location    
* Equipment Blank exceedances  
    + Assessing the influence of field errors using equipment blank samples  
* Parameter Pairs 
    + Flagging samples where component parameters exceed total concentrations
* Holding time exceedances  
    + Flagging samples where holding times were exceeded  

The accuracy, precision, and equipment blank assessments are performed on a subset of samples. The errors discovered with these QC assessments are applied to all normal samples. This is done by associating normal samples to those QC samples closest in date/time to each sample. 

##Formatting Sample data for analysis

```{r, echo=FALSE}
#associate the data with the reporting limts file
rlimits<-read.csv("sections/data/reporting.limits.csv")
data<-merge(data,rlimits,by=c('chemical_name'),all=FALSE)
rm(rlimits)

# Convert dates from factors to POSIXct objects
#library(lubridate)
#if (all(stringr::str_count(data$sample_date, ':') == 2)){
#  data$sample_date <- mdy_hms(data$sample_date)
#  data$analysis_date <- mdy_hms(data$analysis_date)
#} else{
#  data$sample_date <- mdy_hm(data$sample_date)
#  data$analysis_date <- mdy_hm(data$analysis_date)
#}

#pulling spike dataset
spike<-data[data$sample_type_code=="MS",]
spike<-unique(spike[c('sys_sample_code','lab_sdg','sample_date','chemical_name','short','fraction','result_value','result_unit','quantitation_limit','method_detection_limit','detection_limit_unit','lab_qualifiers','qc_original_conc','qc_spike_added','qc_spike_measured','accuracy')])
spike<-spike[!is.na(spike$chemical_name),]

#for all other data sets I need to first restrict the raw data to only that listed as "field"
data<-data[data$sample_source=="Field",]
data<-data[!is.na(data$chemical_name),]

#pulling equipment blank data set
EB<-data[data$DEC_sample_type=="EB",]
EB<-unique(EB[c('sys_sample_code','lab_sdg','sample_date','chemical_name','short','fraction','result_value','result_unit','quantitation_limit','method_detection_limit','detection_limit_unit','lab_qualifiers','reporting_limit','reporting_units')])
EB<-EB[!is.na(EB$chemical_name),]

#pulling duplicate data set
dup<-data[data$DEC_sample_type=="DUP"|data$DEC_sample_type=="N_DUPPARENT",]
dup<-unique(dup[c('DEC_sample_type','sys_sample_code','lab_sdg','sample_date','chemical_name','short','fraction','result_value','result_unit','quantitation_limit','method_detection_limit','detection_limit_unit','lab_qualifiers')])
#separate the parent and duplicate data sets
dupparent<-dup[dup$DEC_sample_type=="N_DUPPARENT",]
dup<-dup[dup$DEC_sample_type=="DUP",]
#remove -DUP from sys sample code for merge
dup$sys_sample_code<-gsub("-DUP","",dup$sys_sample_code)
dup$sys_sample_code<-gsub(" DUP","",dup$sys_sample_code)
dup$sys_sample_code<-gsub("WSDUP","WS",dup$sys_sample_code)

#convert both sys_sample_code fields to character
dup$sys_sample_code<-as.character(dup$sys_sample_code)
dupparent$sys_sample_code<-as.character(dupparent$sys_sample_code)
#rename duplicate fields so can merge the two tables
names(dupparent)[names(dupparent)=='result_value']<-'result_value.parent'
names(dupparent)[names(dupparent)=='result_unit']<-'result_unit.parent'
names(dupparent)[names(dupparent)=='sample_date']<-'sample_date.parent'
names(dupparent)[names(dupparent)=='quantitation_limit']<-'quantitation_limit.parent'
names(dupparent)[names(dupparent)=='method_detection_limit']<-'method_detection_limit.parent'
names(dupparent)[names(dupparent)=='detection_limit_unit']<-'detection_limit_unit.parent'
names(dupparent)[names(dupparent)=='lab_qualifiers']<-'lab_qualifiers.parent'
names(dup)[names(dup)=='result_value']<-'result_value.dup'
names(dup)[names(dup)=='result_unit']<-'result_unit.dup'
names(dup)[names(dup)=='sample_date']<-'sample_date.dup'
names(dup)[names(dup)=='quantitation_limit']<-'quantitation_limit.dup'
names(dup)[names(dup)=='method_detection_limit']<-'method_detection_limit.dup'
names(dup)[names(dup)=='detection_limit_unit']<-'detection_limit_unit.dup'
names(dup)[names(dup)=='lab_qualifiers']<-'lab_qualifiers.dup'
#remove DEC_sample_type for merge
dup<-unique(dup[c('sys_sample_code','lab_sdg','chemical_name','short','sample_date.dup','result_value.dup','result_unit.dup','quantitation_limit.dup','method_detection_limit.dup','detection_limit_unit.dup','lab_qualifiers.dup')])
dupparent<-unique(dupparent[c('sys_sample_code','lab_sdg','chemical_name','short','sample_date.parent','result_value.parent','result_unit.parent','quantitation_limit.parent','method_detection_limit.parent','detection_limit_unit.parent','lab_qualifiers.parent')])
#merge the two into one data set
dup<-merge(dup,dupparent,by=c('sys_sample_code','lab_sdg','chemical_name','short'),all=TRUE)
rm(dupparent)


#remove qc from complete data set
data<-data[data$DEC_sample_type=="N"|data$DEC_sample_type=="N_DUPPARENT",]

#creating a holding time subset
HT<-unique(data[c('sys_sample_code','lab_sdg','sample_date','chemical_name','short','analysis_date')])


```

Plotting the sample dates to understand how best to associate them.  

```{r, echo=FALSE}
samples<-unique(data[c('sample_date','lab_sdg')])
samples$freq<-1
samples$type<-"standard samples"
samples$sample_date<-as.Date(samples$sample_date,"%m/%d/%Y")
blanksamples<-unique(EB[c('sample_date','lab_sdg')])
blanksamples$freq<-2
blanksamples$type<-"equipment blanks"
blanksamples$sample_date<-as.Date(blanksamples$sample_date,"%m/%d/%Y")
dupsamples<-unique(dup[c('sample_date.parent','lab_sdg')])
dupsamples$freq<-3
dupsamples$type<-"duplicates"
names(dupsamples)[names(dupsamples)=="sample_date.parent"]<-"sample_date"
dupsamples$sample_date<-as.Date(dupsamples$sample_date,"%m/%d/%Y")
spikesamples<-unique(spike[c('sample_date','lab_sdg')])
spikesamples$freq<-4
spikesamples$type<-"spike samples"
spikesamples$sample_date<-as.Date(spikesamples$sample_date,"%m/%d/%Y")
#merge together
samples<-merge(samples,blanksamples,all=TRUE)
samples<-merge(samples,dupsamples,all=TRUE)
samples<-merge(samples,spikesamples,all=TRUE)

library(ggplot2)
print(ggplot() +
  geom_point(data=samples,aes(sample_date,freq,color=type)) +
  ylab("sample types")+
  xlab("sample date"))
rm(list=c('samples','blanksamples','dupsamples','spikesamples'))
```

```{r, child = 'sections/Lab.Rmd'}

```



```{r, child = 'sections/Accuracy.Rmd'}

```

```{r, child = 'sections/Precision.Rmd'}

```

```{r, child = 'sections/Equipment.Blanks.Rmd'}

```


```{r, child = 'sections/Parameter.pairs.Rmd'}

```

```{r, child = 'sections/HT.Rmd'}

```

```{r, child = 'sections/Conclusions.Rmd'}

```

```{r, echo=FALSE}
#creating final data set
forprint<-unique(data[c('sys_sample_code','chemical_name','sample_date','cas_rn','fraction','result_value','result_unit','method_detection_limit','detection_limit_unit','quantitation_limit','lab_qualifiers','validator_qualifiers','interpreted_qualifiers')])
```