---
title: "WallkillQAQC"
author: "Alene Onion"
date: "December 29, 2018"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#file.copy("sections/images", ".", recursive = TRUE)
```
 
##QAQC overview

The components of our quality review are:

* Accuracy
    + A matrix spike is used to document the bias of a method in a given sample matrix.   
    + designated 'spike' samples are spiked with a known concentration. The difference between the concentration before and after divided by the spiked amount is the percent recovery. Spike samples are acceptable if the percent recovery is 100% +/- the accuracy criteria established by the standard methods manual and listed in tables 3 and 4.  Otherwise, the spike and it's associated samples are flagged as rejected.
    + NOTE: the QAPP only requires that we analyze spikes for nutrients, metals and minerals (see section II.5.A.b)! I’m presuming this excludes solids, turbidity, conductivity, and hardness in addition to in situ parameters, alkalinity, DOC, and chlorophyll a
    + NOTE: we only examine the matrix spike samples. The matrix spike duplicate samples would only be analyzed when we are meausring organics. That's not the case in any BWAM programs.  
* Precision   
    + Duplicate samples are analyzed from the same date/location to assess the method's precision.    
    + Previously, the % difference between these samples must satisfy the precision criteria set forth in table 3 (screening) and table 4 (routine). In this document, we propose using a new method developed by Ohio EPA that offers a curved threshold.    
* Equipment Blank exceedances  (result > reporting limit)  
    + Step 1: identify a reporting limit based on the data collected in 2017+18
    + Step 2: pull EB values that exceed these reporting limits
    + Step 3: flag sample values that are less than 10x (flagged calculated), 5x (flagged trend), and 3x (rejected)  
* Completeness: Did we collect all the samples we intended to 
* Holding time exceedances  
* Temperature upon receipt exceedances  

##Associating Samples  

Plotting the wallkill sample dates to understand how best to associate them.  

```{r, echo=FALSE}
samples<-unique(wallkill[c('sample_date','lab_sdg')])
samples$freq<-1
samples$type<-"standard samples"
samples$sample_date<-as.Date(samples$sample_date,"%m/%d/%Y")
blanksamples<-unique(EB[c('sample_date','lab_sdg')])
blanksamples$freq<-2
blanksamples$type<-"equipment blanks"
blanksamples$sample_date<-as.Date(blanksamples$sample_date,"%m/%d/%Y")
dupsamples<-unique(dup[c('sample_date.parent','lab_sdg')])
dupsamples$freq<-3
dupsamples$type<-"duplicates"
names(dupsamples)[names(dupsamples)=="sample_date.parent"]<-"sample_date"
dupsamples$sample_date<-as.Date(dupsamples$sample_date,"%m/%d/%Y")
spikesamples<-unique(spike[c('sample_date','lab_sdg')])
spikesamples$freq<-4
spikesamples$type<-"spike samples"
spikesamples$sample_date<-as.Date(spikesamples$sample_date,"%m/%d/%Y")
#merge together
samples<-merge(samples,blanksamples,all=TRUE)
samples<-merge(samples,dupsamples,all=TRUE)
samples<-merge(samples,spikesamples,all=TRUE)

library(ggplot2)
print(ggplot() +
  geom_point(data=samples,aes(sample_date,freq,color=type)) +
  ylab("sample types")+
  xlab("sample date"))
rm(list=c('samples','blanksamples','dupsamples','spikesamples'))
```

## Info from EDD reports
Conclusions drawn from reading lab reports  
I have to read these when I get back. Here are all the SDGs associated with these data:

```{r message=FALSE, warning=FALSE, results='asis'}
unique(wallkill$lab_sdg)
```

```{r, child = 'sections/Accuracy.Rmd'}

```

```{r, child = 'sections/Precision.Rmd'}

```

```{r, child = 'sections/Equipment.Blanks.Rmd'}

```




```{r}
#, }
#child = 'sections/Parameter.pairs.Rmd'}

```

