---
title: "Accuracy"
author: "Alene Onion"
date: "December 29, 2018"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---

```{r setuA, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#file.copy("sections/images", ".", recursive = TRUE)
```

## Accuracy Assessment

* Accuracy
    + A sample matrix spike is used to document the bias of a method in a given sample matrix.   
    + Matrix Spike Samples are collected along with regular water quality samples and spiked in the analytic laboratory with a known concentration of analyte. The samples are then analyzed to determine the accuracy (percent recovery) of the analytic results for a given matrix. Spike samples are acceptable if the percent recovery is 100% +/- the accuracy criteria established by the standard methods manual and listed in tables 3 and 4.  Otherwise, the spike and it's associated samples are flagged as rejected, needs verification. Furthermore, if the sample is >4x the spiked amount, the spike assessment is irrelevant and flagged NA.
    + NOTE: The QAPP only requires that we analyze spikes for nutrients, metals and minerals (see section II.5.A.b). It is presumed that this excludes solids, turbidity, conductivity, and hardness in addition to in situ parameters, alkalinity, DOC, and chlorophyll a.  
    + NOTE: We only examine the matrix spike samples. The matrix spike duplicate samples would only be analyzed when we are meausring organics. That's not the case in any BWAM programs.
    + NOTE: Keep in mind that there may be some parameters where samples would be flagged but they shouldn't be because the spike is diluted even when the sample is <4x the spiked amount.
    + NOTE: Both the laboratory (ALS) and Standards and Analytical Support Section (Rose Ann Garry) recommend using higher thresholds for acceptable %recovery than what is listed in the QAPP. Because of this, we have changed the threshold for all parameters except turbidity and chlorophyll to +/- 25% rather than +/-20%.

The accuracy assessments are performed on a subset of samples. The errors discovered with these QC assessments are applied to all normal samples. This is done by associating normal samples to those QC samples closest in date/time to each sample. 

    
### Accuracy Rejected Samples  
```{r message=FALSE, warning=FALSE, results='asis'}
#calculate the failures myself rather than depending on the lab
spike$percrecovery<-((abs(spike$qc_spike_measured-spike$qc_original_conc))/spike$qc_spike_added)*100
spike$spikeamount<-spike$qc_original_conc/spike$qc_spike_added
spike$pass<-ifelse(spike$spikeamount>4,NA,
            ifelse(spike$percrecover>(100+spike$accuracy)|spike$percrecovery<(100-spike$accuracy),"fail","pass"))

#create a separate file with the samples flagged as pass / fail to merge with parent file later (safety precaution)
#format date to date/time because multiple spike samples in one day
library(lubridate)
# spike$sample_date <- mdy_hm(spike$sample_date)
spikes<-unique(spike[c('chemical_name','sample_date','pass')])

samples<-unique(data[c('sys_sample_code','chemical_name','sample_date')])
# samples$sample_date <- mdy_hm(samples$sample_date)

params<-unique(spikes$chemical_name)
nparams<-length(params)

dataset <- bind_nearest_date(sample.df = samples,
                          match.df = spikes,
                          match.col = "pass") %>% 
  dplyr::rename(spikepass = pass)

#merge with final dataset
#first convert data date to a date
# data$sample_date<-mdy_hm(data$sample_date)
data<-merge(data,dataset,by=c('sys_sample_code','chemical_name','sample_date'),all=TRUE)
data<-data[!is.na(data$chemical_name),]
#convert date back to simple date
# data$sample_date<-as.Date(data$sample_date,"%m/%d/%Y")
rm(dataset)
```

```{r message=FALSE, warning=FALSE, results='asis', fig.height=5.5, fig.width=9}
#plotting the sample failures
#count fails and passes
# library(plyr)
spikefails<-count(spike[c('short','pass')])
samplefails<-count(data[c('short','spikepass')])
names(samplefails)[names(samplefails)=='spikepass']<-'legend'
names(spikefails)[names(spikefails)=='pass']<-'legend'

library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())

 samplefails<-(ggplot() +
    geom_col(data=samplefails,aes(short,freq,fill=legend)) +
      scale_fill_manual(values = c("pass" = "#00BFC4", "fail" = "#F8766D"), na.value = "gray") +
      coord_flip() +
      theme(axis.title.x = element_blank(),axis.title.y = element_blank())) +
      ggtitle("Associated Sample Failures")
 spikefails<-(ggplot() +
    geom_col(data=spikefails,aes(short,freq,fill=legend)) +
      scale_fill_manual(values = c("pass" = "#00BFC4", "fail" = "#F8766D"), na.value = "gray") +
      coord_flip() +
      theme(axis.title.x = element_blank(),axis.title.y = element_blank())) +
      ggtitle("Individual Spike Failures")
 ggarrange(
  samplefails, spikefails,
  common.legend = TRUE, legend = "bottom"
  )
 knitr::kable(spike %>% filter(pass=="fail"))
```

```{r message=FALSE, warning=FALSE, results='asis'}
#This bit of script converts pass faill to R (reject) and A (Accept)
data$spikepass<-gsub("pass","A",data$spikepass)
data$spikepass<-gsub("fail","R",data$spikepass)

```
