---
title: "Conclusions"
author: "Alene Onion"
date: "January 2, 2019"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---

```{r setupC, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#file.copy("sections/images", ".", recursive = TRUE)
```

```{r message=FALSE, warning=FALSE, results='asis'}
```

##Reminder of QAQC component analyses  
The components of our quality review are:

* Laboratory QAQC
    + Checking that laboratory analyses met all internal QAQC checks  
* Accuracy
    + Assessing the bias of a method in a given sample matrix Using a matrix spike   
* Precision   
    + Assessing method precision using duplicate samples analyzed from the same date/location    
* Equipment Blank exceedances  
    + Assessing the influence of field errors using equipment blank samples  
* Parameter Pairs 
    + Flagging samples where component parameters exceed total concentrations
* Holding time exceedances  
    + Flagging samples where holding times were exceeded  
