---
title: "Conclusions"
author: "Alene Onion"
date: "January 2, 2019"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---

```{r setupC, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#file.copy("sections/images", ".", recursive = TRUE)
```

##Summary of QAQC Results

###Reminder of QAQC overview  
The components of our quality review are:

* Laboratory QAQC
    + Checking that laboratory analyses met all internal QAQC checks.  
* Accuracy
    + Assessing the bias of a method in a given sample matrix Using a sample matrix spike.   
* Precision   
    + Assessing method precision using duplicate samples analyzed from the same date/location.    
* Equipment Blank exceedances  
    + Assessing the influence of field errors using equipment blank samples.  
* Parameter Pairs 
    + Flagging samples where component parameters exceed total concentrations.
* Holding time exceedances  
    + Flagging samples where holding times were exceeded.  

The accuracy, precision, and equipment blank assessments are performed on a subset of samples. The errors discovered with these QC assessments are applied to all normal samples. This is done by associating normal samples to those QC samples closest in date/time to each sample. 

```{r message=FALSE, warning=FALSE, results='asis'}
#creating one flag to rule them all and interpetation of this flag

#convert all NA values in pass fields to A
data$labpass<-ifelse(is.na(data$labpass),"A",data$labpass)
data$spikepass<-ifelse(is.na(data$spikepass),"A",data$spikepass)
if (nrow(dup) > 0 ) {
  data$duppass<-ifelse(is.na(data$duppass),"A",data$duppass)
}
data$EBpass<-ifelse(is.na(data$EBpass),"A",data$EBpass)
data$pairpass<-ifelse(is.na(data$pairpass),"A",data$pairpass)
data$HTpass<-ifelse(is.na(data$HTpass),"A",data$HTpass)
#create an overall QA field
data$validator_qualifiers<-"A"
if (nrow(dup) > 0 ) {
  data$validator_qualifiers<-ifelse((data$labpass=="R"|data$spikepass=="R"|data$duppass=="R"|data$EBpass=="R"|data$pairpass=="R"|data$HTpass=="R"),"R",
             ifelse(data$EBpass=="T","T",
             ifelse((data$labpass=="J"|data$EBpass=="J"),"J",
             ifelse(data$labpass=="U","U",data$validator_qualifiers))))
} else{
    data$validator_qualifiers<-ifelse((data$labpass=="R"|data$spikepass=="R"|data$EBpass=="R"|data$pairpass=="R"|data$HTpass=="R"),"R",
             ifelse(data$EBpass=="T","T",
             ifelse((data$labpass=="J"|data$EBpass=="J"),"J",
             ifelse(data$labpass=="U","U",data$validator_qualifiers))))
}
#add explanation for flagged data for each type
#first create the field
data$interpreted_qualifiers<-"Normal"

#For lab errors
#R flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$labpass=="R","lab error","Normal"),
                                ifelse(data$labpass=="R",paste(data$interpreted_qualifiers," ; lab error",sep=""),data$interpreted_qualifiers))
#J flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$labpass=="J","estimated value due to lab error","Normal"),
                                ifelse(data$labpass=="J",paste(data$interpreted_qualifiers," ; estimated value due to lab error",sep=""),data$interpreted_qualifiers))
#U flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$labpass=="U","Analyte was analyzed for but not detected","Normal"),
                                ifelse(data$labpass=="U",paste(data$interpreted_qualifiers," ; Analyte was analyzed for but not detected",sep=""),data$interpreted_qualifiers))

#For EB errors
#R flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$EBpass=="R","Equipment Blank error","Normal"),
                                ifelse(data$EBpass=="R",paste(data$interpreted_qualifiers," ; Equipment Blank error",sep=""),data$interpreted_qualifiers))
#J flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$EBpass=="J","estimated value due to Equipment Blank error","Normal"),
                                ifelse(data$EBpass=="J",paste(data$interpreted_qualifiers," ; estimated value due to Equipment Blank error",sep=""),data$interpreted_qualifiers))

#For spike errors
#R flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$spikepass=="R","Accuracy error","Normal"),
                                ifelse(data$spikepass=="R",paste(data$interpreted_qualifiers," ; Accuracy error",sep=""),data$interpreted_qualifiers))

#For dup errors
#R flags
if (nrow(dup) > 0 ) {
  data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                  ifelse(data$duppass=="R","Precision error","Normal"),
                                  ifelse(data$duppass=="R",paste(data$interpreted_qualifiers," ; Precision error",sep=""),data$interpreted_qualifiers))
}

#For pair errors
#R flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$pairpass=="R","Result should be less than the total","Normal"),
                                ifelse(data$pairpass=="R",paste(data$interpreted_qualifiers," ; Result should be less than the total",sep=""),data$interpreted_qualifiers))

#For HT errors
#R flags
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",
                                ifelse(data$HTpass=="R","Holding Time error","Normal"),
                                ifelse(data$HTpass=="R",paste(data$interpreted_qualifiers," ; Holding Time error",sep=""),data$interpreted_qualifiers))

#changing the 'Normal' interpreted flag to ""
data$interpreted_qualifiers<-ifelse(data$interpreted_qualifiers=="Normal",'',data$interpreted_qualifiers)

```

###Flagged Samples  

Data flags applied:  

```{r message=FALSE, warning=FALSE, results='asis'}
flags<-read.csv("data/validator_flags.csv")

knitr::kable(flags)
rm(flags)
```    
  
```{r message=FALSE, warning=FALSE, results='asis', fig.height=5.5}
#plotting the sample failures
#count fails and passes
# library(plyr)
labfails<-count(data[c('short','validator_qualifiers')])
names(labfails)[names(labfails)=='validator_qualifiers']<-'legend'

library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())

 print(labfails<-(ggplot() +
    geom_col(data=labfails,aes(short,freq,fill=legend)) +
      scale_fill_manual(values = c("A" = "#00BFC4", "J" = "#C77CFF", "U" = "gray", "R" = "#F8766D", "T" = "#7CAE00")) +
      coord_flip() +
      theme(axis.title.x = element_blank(),axis.title.y = element_blank()))) 
# rm(list=c('labfails'))
```

###Interpreted Flags  
Plotting the interpreted flags to see the reason the data were flagged:  
```{r message=FALSE, warning=FALSE, results='asis', fig.height=5.5}
#plot the result value / day for each param 
#color the points based on flags
labfails<-unique(data$chemical_name)
labfails<-count(data[c('short','interpreted_qualifiers')])
names(labfails)[names(labfails)=='interpreted_qualifiers']<-'legend'


library(ggplot2)
library(ggpubr)
library(pals)
theme_set(theme_pubr())


 #plot the graph without the legend
 print((ggplot() +
    geom_col(data=labfails,aes(short,freq,fill=legend)) +
      scale_fill_manual(values= c("gray", (as.vector(alphabet2(19))))) +
      coord_flip() +
      theme(axis.title.x = element_blank(),axis.title.y = element_blank(),legend.position="none"))) 
#plot just the legend
 legend<-(ggplot() +
                  geom_col(data=labfails,aes(short,freq,fill=legend)) +
                  coord_flip() +
                  theme_void()) +
         scale_fill_manual(values= c("gray", (as.vector(alphabet2(19)))))
legend<-get_legend(legend)
as_ggplot(legend)
rm(legend)

rm(list=c('labfails'))

```



